\documentclass{SelimArticle}
\input{./options}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\pdiff}[2]{\ensuremath{
    \frac{\partial #1}{\partial #2}
}}
\newcommand{\ip}{\ensuremath{_{i+1,j}}}
\newcommand{\im}{\ensuremath{_{i-1,j}}}
\newcommand{\jp}{\ensuremath{_{i,j+1}}}
\newcommand{\jm}{\ensuremath{_{i,j-1}}}
\begin{document}
\mytitlepage
% \tableofcontents
\newpage
% Begin writing here.
\section{Governing Differential Equation}
The Laplace equation in two dimensions is given by:
\begin{equation}
    \label{eq:laplace}
    \frac{\partial^2 u}{\partial x^2} + \frac{\partial^2 u}{\partial y^2} = 0
\end{equation}
The boundary conditions are:
\begin{gather*}
    u(x,0) = u(0,y) = u(1,y) = 0\\
    u(x,1) = 1
\end{gather*}

\section{Question 1}
\begin{quote}
    \textit{
     Derive the truncation error for the second-order finite-difference spatial discretization
     for the Laplace equation.}
\end{quote}
A second-order finite-difference spatial discretization for a second derivative is given by:
\begin{equation}
    \label{eq:secord}
    \pdiff{^2u}{x^2} = \frac{u\ip^n - 2 u_{ij}^n + u\im^n}{\Delta x^2}
\end{equation}
Thus, plugging~\Cref{eq:secord} into~\Cref{eq:laplace} yields the following discretization:
\begin{equation}
    \label{eq:laplacediscrete}
    \frac{u\ip^n - 2 u_{ij}^n + u\im^n}{\Delta x^2}
    + \frac{u\jp^n - 2 u_{ij}^n + u\jm^n}{\Delta y^2} = 0
\end{equation}
Now, we substitute each term by its Taylor series expansion into~\Cref{eq:laplacediscrete} to obtain
the modified equation:
\begin{equation*}
    \label{eq:te}
    u_{xx} + u_{yy} =
    \underbrace{
        - \frac{1}{12}\left(u_{xxxx}\Delta x ^2 + u_{yyyy}\Delta y^2 \right)
        + O(\Delta x^4, \Delta y^4)
    }_{\text{Truncation Error}
\end{equation*}
Note the leading error terms are a function of $\Delta x^2, \Delta y^2$, hence the scheme being
second-order accurate.
\section{Question 2}
\begin{quote}
    \textit{Derive the modified equation for the scheme and note whether the error is dissipative,
    dispersize, or a combination of the two. Explain why?}
\end{quote}
As derived in the previous question, the truncation error is given by:
$$
    \frac{1}{12}\left(u_{xxxx}\Delta x ^2 + u_{yyyy}\Delta y^2 \right)
    + O(\Delta x^4, \Delta y^4)
$$
Due to the discretization, all the odd-order derivatives vanish, leaving only the even-order ones.
Now, we know that a second-order derivative in the truncation error introduces dissipation.
The same can be said of all even-order derivatives.

Thus, the chosen scheme induces dissipative errors \textbf{only}.

\section{Question 3}
\begin{quote}
    \textit{Demonstrate the solution of the Laplace Equation for the 400 × 400.}
\end{quote}

\end{document}
